Predicting Transport in Lyotropic Liquid Crystal Membranes with Molecular Dynamics Simulations

### 2017-08-17 Ben_Manuscripts/Masters_Bypass/methods.tex
updates

Initial guesses for the remaining structural parameters were chosen
based on experimental data and treated as variables during model
development. The distance between pores was based on experimental SAXS
data for this system \cite{feng_thin_2016}. Initial configuration pore
spacing was chosen to be 10 \% larger than the experimental value to reduce
unintended repulsions resulting from a tightly packed initial configuration. 
The layer spacing was based on experimental 2D WAXS data which shows
reflections corresponding to features spaced 3.7 \angstrom apart. It has 
been hypothesized that the features are present due to pi-pi
interactions between stacked aromatic rings \cite{feng_scalable_2014}. 
Our simulations tend to equilibrate to a wider interlayer spacing of $\approx$ 4.1 \angstrom,
which inspired separate systems starting with layer spacings greater than
 4 \angstrom. Initial pore radius estimates of 0.6 nm were based on past TEM 
images and size exclusion rejection data\cite{feng_scalable_2014,
feng_thin_2016,zhou_supported_2005}.
% There is plenty I can say about initial configuration. Starting layers
% too far apart, too close, tilted initial configurations etc.
The relative interlayer orientation was chosen based on clues from diffraction
data as well as the various known stacking modes of benzene and substituted 
benzene rings: sandwiched, parallel-displaced and T-shaped 
(Figures~\ref{fig:sandwiched,fig:pd,fig:tshaped}). The T-shaped 
configuration was ruled out based on the inconsistency of its $\approx$ 
5 \angstrom equilibrium stacking distance. The system's preference
towards the sandwiched vs. parallel displaced stacking modes was left
to be explored. Visualization of each configuration 
(Figures~\ref{fig:sandwiched_layers,fig:offset_layers} suggests 
entropic differences based on the way the tails are able to pack. 
This difference may explain, in part, which stacking mode is more favorable.

\begin{figure}
	\centering
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=\textwidth]{sandwiched.png}
		\caption{}\label{fig:sandwiched}
	\end{subfigure}
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=\textwidth]{PD.png}
		\caption{}\label{fig:pd}
	\end{subfigure}
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Tshaped.png}
		\caption{}\label{fig:tshaped}
	\end{subfigure}
	\vskip\baselineskip
	\begin{subfigure}[b]{0.475\textwidth}
		\centering
		\includegraphics[width=\textwidth]{sandwiched_layers.png}
		\caption{}\label{fig:sandwiched_layers}
	\end{subfigure}
	\begin{subfigure}[b]{0.475\textwidth}
		\centering
		\includegraphics[width=\textwidth]{offset_layers.png}
		\caption{}\label{fig:offset_layers}
	\end{subfigure}
	\caption{}\label{fig:stacking}
\end{figure}

>>>>

Initial guesses for the remaining structural parameters were chosen
based on experimental data and treated as variables during model
development. The distance between pores was based on experimental SAXS
data for this system \cite{feng_thin_2016}. Initial configuration pore
spacing was chosen to be 10 \% larger than the experimental value to reduce
unintended repulsions resulting from a tightly packed initial configuration. 
The layer spacing was based on experimental 2D WAXS data which shows
reflections corresponding to features spaced 3.7 \angstrom apart. It has 
been hypothesized that the features are present due to pi-pi
interactions between stacked aromatic rings \cite{feng_scalable_2014}. 
Our simulations tend to equilibrate to a wider interlayer spacing of $\approx$ 4.1 \angstrom,
which inspired separate systems starting with layer spacings greater than
 4 \angstrom. Initial pore radius estimates of 0.6 nm were based on past TEM 
images and size exclusion rejection data\cite{feng_scalable_2014,
feng_thin_2016,zhou_supported_2005}.
% There is plenty I can say about initial configuration. Starting layers
% too far apart, too close, tilted initial configurations etc.
The relative interlayer orientation was chosen based on clues from diffraction
data as well as the various known stacking modes of benzene and substituted 
benzene rings: sandwiched, parallel-displaced and T-shaped 
(\Cref{fig:sandwiched,fig:pd,fig:tshaped}). The T-shaped 
configuration was ruled out based on the inconsistency of its $\approx$ 
5 \angstrom equilibrium stacking distance. The system's preference
towards the sandwiched vs. parallel displaced stacking modes was left
to be explored. Visualization of each configuration 
(\Cref{fig:sandwichedlayers,fig:offsetlayers}) suggests 
entropic differences based on the way the tails are able to pack. 
This difference may explain, in part, which stacking mode is more favorable.

\begin{figure}
	\centering
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=\textwidth]{sandwiched.png}
		\caption{Sandwiched}\label{fig:sandwiched}
	\end{subfigure}
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=\textwidth]{PD.png}
		\caption{Parallel-Displaced}\label{fig:pd}
	\end{subfigure}
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Tshaped.png}
		\caption{T-shaped}\label{fig:tshaped}
	\end{subfigure}
	\vskip\baselineskip
	\begin{subfigure}[b]{0.475\textwidth}
		\centering
		\includegraphics[width=\textwidth]{sandwichedlayers.png}
		\caption{Two monomer layers stacked in the sandwiched configuration}\label{fig:sandwichedlayers}
	\end{subfigure}
	\begin{subfigure}[b]{0.475\textwidth}
		\centering
		\includegraphics[width=\textwidth]{offsetlayers.png}
		\caption{Two monomer layers stacked in the parallel-displaced configuration}\label{fig:offsetlayers}
	\end{subfigure}
	\caption{}\label{fig:stacking}
\end{figure}

### 2017-08-17 Ben_Manuscripts/Masters_Bypass/results.tex
revisions
updates

To understand the composition of the monomer layers, we ran simulations of 
systems created with 4 - 8 monomers per layer. Both the sandwiched and 
parallel displaced configurations were tested. All systems remained
stable for at least 50 ns. Pair correlation functions calculated between 
aromatic rings along the length of the pores indicate that systems prefer to
stay partitioned into layers (Fig~\ref{fig:zdf}). Pore spacings were 
calculated by measuring the distance between the pore centers. Pore centers
were located by averaging the coordinates of sodium ions in their 
respective pores. Statistics were generated using the bootstrapping 
technique. Systems built with 5 monomers in each layer equilibrate to a pore
spacing that is most consistent with the experimental value of 4.12 nm
derived from SAXS measurements (Table~\ref{table:p2p}). The remainder of
this discussion will focus on the analysis of systems built with 5 
monomers per layer.

The experimental 2D WAXS measurement contains five major reflections which
our model intends to reproduce (Fig ~\ref{fig:WAXSexp}). The first is located
at qz = 1.7 \angstrom \textsuperscript{-1}, corresponding to a real space 
spacing of 3.7 \angstrom. The reflection is attributed to pi-pi stacking
between aromatic rings in the direction perpendicular to the membrane plane, or z-axis. 
For simplicity, this reflection will be referred to as R-pi. A weak intensity
line is located at exactly half the qz value of R-pi (qz = 0.85 \angstrom \textsuperscript{-1}), corresponding to a 
real space periodic spacing of 7.4 \angstrom. This reflection has been 
attributed to 2\textsubscript{1} helical ordering of aromatic rings along the
z axis and will therefore be referred to as R-helix. If positions of the 
aromatic rings can be traced by a helix, then for each turn in the helix, 
there should be two aromatic rings. A third major reflection is marked by 
a low intensity ring located at r = 1.4 \angstrom \textsuperscript{-1}. The
real space separation corresponds to 4.5 \angstrom which is characteristic
of the average spacing between packed alkane chains. This reflection will be
called R-alkanes. Within R-alkanes, are four spots of higher intensity which 
will be called R-spots. All are located $\approx 40$ degrees from the qz axis
in their respective quadrants. In many liquid crystal systems this can be
attributed to the tilt angle of the alkane chains with respect to the xy axis. 
The final major reflections correspond to the spacing and symmetry of the d100 
plane which can be related to the distance between pores. The feature, which 
will be called R-pores, is characterized by dots along qz = 0. The spacing
between dots is indicative of the hexagonal symmetry of the packed pores. It
is easiest to interpret the data by radially integrating the 2D data to get 
a 1D curve which has also been done.

\begin{figure}
	\centering
	\begin{subfigure}{0.47\linewidth}
		\centering
		\includegraphics[width=\linewidth, valign=t]{WAXS_soft_confined.png}
		\caption{}\label{fig:WAXS}
	\end{subfigure}
	\begin{subfigure}{0.43\linewidth}
		\centering
		\vspace{12mm}
		\includegraphics[width=\linewidth]{SAXS.png}
		\caption{}\label{figs:SAXS}
	\end{subfigure}
	\caption{(a) 2D wide angle X-ray scattering gives details about repeating features less than 1 nanometer apart (b) 1D small angle X-ray scattering indicates hexagonal packing of pores as well as the spacing between pores.}\label{fig:SWAXS}
\end{figure}

The initial distance between layers can influence the approach towards
equilibrium. Equilibrated systems built according to the 3.7 \angstrom
spacing indicated by R-pi are characterized by a defined, cylindrical and
open pore structure. Aromatic rings are arranged in a helical conformation.
We will refer to this system as Basin A (Fig ~\ref{fig:phaseA}). Simulations
of systems built with layers stacked > 4 \angstrom apart results in a pore
structure characterized by high radial disorder, while still maintaining 
partitioning between hydrophobic and hydrophilic regions. This will be 
called Basin B (Fig~\ref{fig:phaseB}). It is apparent that this LLC membrane
may exist in at least two metastable states. The distinct difference in 
pore structure (Fig~\ref{fig:porestructures}) exhibited by each phase will
likely lead to different transport mechanisms. Understanding which phase 
exists experimentally is necessary in order to appropriately study the system.
% BJC: include more examples of the difference between the two basins
% The arrangement of sodium ions (which are closely bound to carbonyl
% head groups) can be well-approximated by a gaussian distribution 
% Like phase A, phase B can form at 280K. 
% The phase is also present when phase A is heated close to its isotropic transition point
% There are distinct differences in the membrane and pore structures between each state (Fig~\ref{fig:porestructures})

We varied the relative interlayer orientation between sandwiched and 
parallel-displaced based on our knowledge of the stability of these two
pi-stacking modes. Short NVT simulations were run with position restraints
on aromatic ring carbon atoms. Simulated X-ray diffraction patterns 
generated from each configuration establish a difference between the two
stacking modes. In each pattern, R-alkanes is present at a distance of
$\approx 1.5$ \angstrom \textsuperscript{-1}. R-pi is also present in each 
pattern, although it appears to intersect R-alkanes because the spacing
between aromatic rings is similar to the alkane chain packing. The difference 
in aromatic ring spacing between experiment and simulation is likely a result
of the inability of GAFF to properly handle pi interactions. The parallel
displaced configuration contains two lines of high intensity with maximum
intensity occurring where they intersect the alkane chain region due to
constructive interference between the two features. The lines are located
where one would expect to see R-helix. As the simulation is progressed, 
the full line begins to fade and the reflection persists only at the 
intersection with the R-alkanes. Most significantly, the line is no
longer present where R-helix exists and, perhaps coincidentally, the 
intersection with R-alkanes occurs where one would expect to see R-spots.

Full comparison of experimental 2D WAXS with simulated X-ray diffraction
patterns produced from equilibrated MD trajectories shows the most consistency
with the sandwiched conformation in Basin A. Systems were built in both the 
parallel displaced and sandwiched conformations with an intial layer spacing of
3.7 \angstrom. A third system was created by stacking layers in the sandwiched 
conformation 5 \angstrom apart in order to guide it towards Basin B. For ease of
reference, the third system will be referred to as the disordered system. The 
three systems were equilibrated according to our procedure with NPT simulations of 
greater than 400 ns. Simulated diffration patterns were generated using 
portions of the trajectory after equilibration. Equilibration was detected
when the distance between pores and the membrane thickness stopped changing.
Simulated diffraction patterns for all three structures are shown in 
Fig~\ref{fig:XRDsim}. The equilibrated parallel displaced structure exhibits 
R-alkanes and R-pores in the expected locations. Due to low resolution, making
out the individual spots of R-pores is challenging and can be validated upon
further inspection of the 3D structure factor. R-pi is also present, however it
intersects R-alkanes because it is at a lower q than expected. This can be 
explained by our model's inability to appropriately represent aromatic interactions.
The rings prefer to stack $\approx 4.1 \angstrom$ apart as opposed to 3.7 \angstrom.
R-helix and R-spots are both missing. The line which intersected both of these
features in the restrained simulations faded upon equilibration which suggests
that monomers in the parallel-displaced conformation may not give rise to the 
experimental WAXS pattern seen. Simulated XRD of the sandwiched configuration 
contains all features except for R-helix. Most notably, R-spots appears in the
expected location, which suggests that there is something intrinsically to
the sandwiched packing which gives rise to such features. Finally, the 
disordered structure exhibits R-alkanes and R-pores. R-helix are R-spots are not
present. Given the evidence, the sandwiched configuration with layers spaced
3.7 \angstrom apart gives rise to an equilibrated structure closest to experiment.

@@

>>>>

To discern the composition of the monomer layers, we ran simulations of 
systems created with 4 - 8 monomers per layer. Both the sandwiched and 
parallel displaced configurations were tested. All systems remained
stable for at least 50 ns. Pair correlation functions calculated between 
aromatic rings along the length of the pores indicate that systems prefer to
stay partitioned into layers (Fig~\ref{fig:zdf}). Pore spacings were 
calculated by measuring the distance between the pore centers. Pore centers
were located by averaging the coordinates of sodium ions in their 
respective pores. Statistics were generated using the bootstrapping 
technique. Table~\ref{table:p2p} shows the pore spacing for all systems tested.
Systems built with 5 monomers in each layer equilibrate to a pore
spacing that is most consistent with the experimental value of 4.12 nm
derived from SAXS measurements (Figure~\ref{fig:SAXS}). The remainder of
this discussion will focus on the analysis of systems built with 5 
monomers per layer.

The experimental 2D WAXS measurement contains five major reflections which
our model intends to reproduce (Figure~\ref{fig:WAXS}). The first is located
at qz = 1.7 \angstrom \textsuperscript{-1}, corresponding to a real space 
spacing of 3.7 \angstrom. The reflection is attributed to pi-pi stacking
between aromatic rings in the direction perpendicular to the membrane plane, or z-axis. 
For simplicity, this reflection will be referred to as R-pi. A weak intensity
line is located at exactly half the qz value of R-pi (qz = 0.85 \angstrom \textsuperscript{-1}), corresponding to a 
real space periodic spacing of 7.4 \angstrom. This reflection has been 
interpreted  as 2\textsubscript{1} helical ordering of aromatic rings along the
z axis and will therefore be referred to as R-helix. If positions of the 
aromatic rings can be traced by a helix, then for each turn in the helix, 
there should be two aromatic rings. A third major reflection is marked by 
a low intensity ring located at r = 1.4 \angstrom \textsuperscript{-1}. The
real space separation corresponds to 4.5 \angstrom which is characteristic
of the average spacing between packed alkane chains. This reflection will be
called R-alkanes. Within R-alkanes, are four spots of higher intensity which 
will be called R-spots. All are located $\approx 40$ degrees from the qz axis
in their respective quadrants. In many liquid crystal systems this can be
explained by the tilt angle of the alkane chains with respect to the xy plane. 
The final major reflections correspond to the spacing and symmetry of the d100 
plane which can be related to the distance between pores. The feature, which 
will be called R-pores, is characterized by dots along qz = 0. The spacing
between dots is indicative of the hexagonal symmetry of the packed pores. It
is easiest to interpret the data by radially integrating the 2D data to get 
a 1D curve which is shown in Figure~\ref{fig:SAXS}.

\begin{figure}
	\centering
	\begin{subfigure}[t]{0.47\linewidth}
		\centering
		\raisebox{.2\textwidth}{%
		\includegraphics[width=\linewidth]{WAXS_soft_confined.png}
		}
		\caption{}\label{fig:WAXS}
	\end{subfigure}
	\begin{subfigure}[t]{0.43\linewidth}
		\centering
	%	\vspace{12mm}
		\includegraphics[width=\linewidth]{SAXS.png}
		\caption{}\label{fig:SAXS}
	\end{subfigure}
	\caption{(a) 2D wide angle X-ray scattering gives details about repeating features less than 1 nanometer apart (b) 1D small angle X-ray scattering indicates hexagonal packing of pores as well as the spacing between pores.}\label{fig:SWAXS}
\end{figure}

The initial distance between layers can influence the approach towards
equilibrium. Equilibrated systems built according to the 3.7 \angstrom 
spacing indicated by R-pi are characterized by a defined, cylindrical and
open pore structure. Aromatic rings are arranged in a helical conformation.
We will refer to this system as Basin A (Figure~\ref{fig:280K_pore}). Simulations
of systems built with layers stacked greater than 4 \angstrom apart results in a pore
structure characterized by high radial disorder, while still maintaining 
partitioning between hydrophobic and hydrophilic regions. This will be 
called Basin B (Figure~\ref{fig:340K_pore}). It is apparent that this LLC membrane
may exist in at least two metastable states. The distinct difference in 
pore structure exhibited by each phase will likely lead to different transport
mechanisms. Understanding which phase exists experimentally is necessary in 
order to appropriately study the system.
% BJC: include more examples of the difference between the two basins
% The arrangement of sodium ions (which are closely bound to carbonyl
% head groups) can be well-approximated by a gaussian distribution 
% Like phase A, phase B can form at 280K. 
% The phase is also present when phase A is heated close to its isotropic transition point
% There are distinct differences in the membrane and pore structures between each state (Fig~\ref{fig:porestructures})

Basin A and Basin B are two configurationally metastable basins. 
We hesitate to classify the basins as separate phases. We hypothesize
that Basin B is present at higher temperatures. This is evidenced by
a simulation of a system in Basin A whose temperature was linearly
ramped from 280K to 340K over 100 ns. Visually, there is a distinct
change in pore structure from one characteristic of Basin A 
(Figure~\ref{fig:280K_pore} to one characteristic of Basin B 
(Figure~\ref{fig:340K_pore}. To quantify our observation, we chose 
three easily measurable order parameters: the distance between pores,
the membrane the thickness and the ratio of pore radius to the uncertainty
in pore radius. The distance between pores was measured as described 
previously. The membrane thickness was measured using the length of the
z box vector. The pore radius was measured by calculating the distance
between the center of mass of each aromatic ring and the center of mass
of all aromatic rings in respective pores. The pore radius is divided by
its uncertainty in order to understand the degree to which monomers 
obstruct the pore region. Referring back to our temperature ramped system,
the slope of all order parameters change around 310K indicating the 
possibility of an abrupt change in system ordering. To prove that the 
basins are in fact two different phases, we must observe evidence of a
first order phase transition, which would be indicated by a discontinuity
in an order parameter as a function of temperature. Our 100 ns temperature
ramp may have been too fast and may have suffered hysteresis. To mitigate
hysteresis, we performed a slow, stepwise temperature ranp. Started from
structures equilibrated at 280K, the system temperature was raised to 
300K and allowed to equilibrate for 200 ns. At the end of 200 ns, the
temperature was raised 5K and equilibrated again for 200 ns. We repeated
this process until the system was equilibrated for 200 ns at 325K. In all
cases, we observed smooth changes in order parameters as temperature was
increased implying that we have not found two phases.

\begin{figure}
        \centering
        \begin{subfigure}[b]{0.45\textwidth}
                \centering
                \includegraphics[width=\textwidth]{280K_tramp_close.png}
                \caption{}\label{fig:280K_pore}
        \end{subfigure}
        \begin{subfigure}[b]{0.45\textwidth}
                \centering
                \includegraphics[width=\textwidth]{340K_tramp_close.png}
                \caption{}\label{fig:340K_pore}
        \end{subfigure}
        \vskip\baselineskip
        \begin{subfigure}[b]{0.325\textwidth}
                \centering
                \includegraphics[width=\textwidth]{p2p_tramp.png}
                \caption{}\label{fig:p2p_tramp}
        \end{subfigure}
        \begin{subfigure}[b]{0.325\textwidth}
                \centering
                \includegraphics[width=\textwidth]{thickness_tramp.png}

                \caption{}\label{fig:thickness_tramp}
        \end{subfigure}
        \begin{subfigure}[b]{0.325\textwidth}
                \centering
                \includegraphics[width=\textwidth]{order_tramp.png}
                \caption{}\label{fig:order_tramp}
        \end{subfigure}
        \caption{}\label{fig:tramp}
\end{figure}

We varied the relative interlayer orientation between sandwiched and 
parallel-displaced based on our knowledge of the stability of these two
pi-stacking modes. Short NVT simulations were run with position restraints
on aromatic ring carbon atoms. Simulated X-ray diffraction patterns 
generated from each configuration establish a difference between the two
stacking modes. In each pattern, R-alkanes is present at a distance of
$\approx 1.5$ \angstrom \textsuperscript{-1}. R-pi is also present in each 
pattern, although it appears to intersect R-alkanes because the spacing
between aromatic rings is similar to the alkane chain packing. The difference 
in aromatic ring spacing between experiment and simulation is likely a result
of the inability of GAFF to properly handle aromatic interactions. The parallel
displaced configuration contains two lines of high intensity with maximum
intensity occurring where they intersect the alkane chain region due to
constructive interference between the two features. The lines are located
where one would expect to see R-helix. As the simulation is progressed, 
the full line begins to fade and the reflection persists only at the 
intersection with the R-alkanes. Most significantly, the line is no
longer present where R-helix exists and, perhaps coincidentally, the 
intersection with R-alkanes occurs where one would expect to see R-spots.

Full comparison of experimental 2D WAXS with simulated X-ray diffraction
patterns produced from equilibrated MD trajectories shows the most consistency
with the sandwiched conformation in Basin A. Systems were built in both the 
parallel displaced and sandwiched conformations with an intial layer spacing of
3.7 \angstrom. A third system was created by stacking layers in the sandwiched 
conformation 5 \angstrom apart in order to guide it towards Basin B. For ease of
reference, the third system will be referred to as the disordered system. The 
three systems were equilibrated according to our procedure with NPT simulations of 
greater than 400 ns. Simulated diffration patterns were generated using 
portions of the trajectory after equilibration. Equilibration was detected
when the distance between pores and the membrane thickness stopped changing.
Simulated diffraction patterns for all three structures are shown in 
Fig~\ref{fig:XRDsim}. The equilibrated parallel displaced structure exhibits 
R-alkanes and R-pores in the expected locations. Due to low resolution, making
out the individual spots of R-pores is challenging and can be validated upon
further inspection of the 3D structure factor. R-pi is also present, however it
intersects R-alkanes because it is at a lower q than expected. This can be 
explained by our model's inability to appropriately represent aromatic interactions.
The rings prefer to stack $\approx 4.1 \angstrom$ apart as opposed to 3.7 \angstrom.
R-helix and R-spots are both missing. The line which intersected both of these
features in the restrained simulations faded upon equilibration which suggests
that monomers in the parallel-displaced conformation may not give rise to the 
experimental WAXS pattern seen. Simulated XRD of the sandwiched configuration 
contains all features except for R-helix. Most notably, R-spots appears in the
expected location, which suggests that there is something intrinsic to
the sandwiched packing which gives rise to such features. Finally, the 
disordered structure exhibits R-alkanes and R-pores. R-helix and R-spots are not
present. Given the evidence, the sandwiched configuration with layers spaced
3.7 \angstrom apart gives rise to an equilibrated structure closest to experiment.

@@

\begin{wrapfigure}{R}{0.4\textwidth}
	\centering
	\includegraphics[width=0.4\textwidth]{tilt.png}
	\caption{The average angle between alkyl chains and the xy plane is nearly zero degrees}\label{fig:tilt}
\end{wrapfigure}
